Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-05 mr-review.Rmd
carriage return
minor addns

Methods for Mendelian randomization have appeared at a rapid and accelerating pace since the 
appearance of the first GWAS (Figure \@ref(fig:article-counts)).

>>>>

Methods for and uses of MR have appeared at a rapid and accelerating pace since the 
appearance of the early GWAS (Figure \@ref(fig:article-counts)). Inverse variance weighted MR [@burgess_mendelian_2013], MR-Egger [@bowden_mendelian_2015], weighted median estimation [@bowden_consistent_2016], Bayesian weighted MR [@zhao_bayesian_2019], robust adjusted profile score [@zhao_statistical_2020], MRMix [@qi_mendelian_2019], CAUSE [@morrison_mendelian_2020], and MRAID [@yuan_likelihood_2021] are all recent methods developments. These methods differ in their approaches to two challenges:

1. instrument selection (from among the available SNPs)  
1. modeling and controlling for horizontal pleiotropy

Many MR methods use a set of pre-selected SNPs as instruments. Typically, these SNPs are selected to 
be statistically independent. This restriction to 
independent SNPs is needed for valid inference in methods like inverse variance weighting MR [@burgess_mendelian_2013]. The independent SNPs may be 
selected through linkage disequilibrium (LD) clumping. In this procedure, SNPs 
are ranked by the strength of their marginal association with the exposure 
variable. The LD clumping procedure then selects SNPs that are not in high LD 
with the top-ranking SNPs.

LD clumping for choosing instrument SNPs can be problematic for several 
reasons. First, the chosen SNPs may not be causal for the outcome. This can 
happen, for instance, by choosing a noncausal SNP that is in high LD with a 
causal SNP.

@verbanck_detection_2018
